---
title: "Alannah's R Notebook"
output: html_notebook
---

Before starting the script, put these files on your Desktop:
1) RTestData.txt
2) primer.csv
3) r_bash.sh r.py test.r

#READING DATAFILES & SETTING VARIABLES

#First start R, then run these commands


```{r}
list.files('~') #are all the files present? 
#Yes!
d1=read.table('RTestData.txt', header=TRUE)
d1
fuso
d1$fuso
attach(d1)
fuso
```

#Plot d1$strep using the hist function.
```{r}
hist(d1$strep)
```

#Run qqnorm with d1$strep - is it linear?
#somewhat.
```{r}
qqnorm(d1$strep)
```

#Do a log transform of d1$strep
```{r}
lstrep=log(d1$strep)
print(lstrep)
```


#Run qqnorm of lstrep
```{r}
qqnorm(lstrep)
```

Plot lstrep with the hist function
```{r} 
hist(lstrep)
```{r}

```

```

#Run a t.test with d$strep and d$fuso
```{r}
t.test(d1$strep, d1$fuso) #values of: t = 3.3584, df = 97.994, p-value = 0.001117
```


#Run a t.test with lstrep and d$fuso
```{r}
t.test(lstrep, d1$fuso) #values of: t = -10.219, df = 74.469, p-value = 8.04e-16
```


#(How did the test statistic and p-value change? Why did the p-value change?)
#For the second t-test, the p-value became much smaller. This is due to the difference in means and higher degrees of freedom.


#Put in the same directory as your R script
#show the data frame, and show the v1 variable.
```{r}
d2=read.csv("primer-1.csv", header=='TRUE')
attach(d2)
d2
d2$v1
```

#SORTING SUBSETTING DATA
```{r}
newdata=d1[order(time),]
time1=subset(d1, time==1)
time1
```

#https://www.statmethods.net/management/subset.html

#MORE COMPLEX DATA WRANGLING - Make a bookmark of this website. 
#https://r4ds.had.co.nz/transform.html


#WRITING FILES
```{r}
write.csv(time1,"time1.csv") #Put in the same directory as your R script
```


#RUN ON THE COMMAND LINE
Using the terminal, run the following scripts. The scripts may need altering or permissions changes.
#okay!

#r_bash.sh 
#r.py


#R Lab 1!
#Use R to perform analyses on two datasets. Use the R Cookbook to figure out how to do all the
analyses.
#Datasets to analyze (R built in datasets)
#(1) airquality
#(2) ToothGrowth

#1. Show the top of the dataset using head
```{r}
head(airquality)
head(ToothGrowth)
```

#2. Attach variables to be easier to analyze: attach(dataframe)
```{r}
attach(airquality)
attach(ToothGrowth)
```

#3. Use 3 different methods to determine normality: qqnorm, a histogram, stats test
```{r}
qqnorm(airquality$Ozone)
hist(airquality$Wind)
t.test(airquality$Temp)
```
```{r}
qqnorm(ToothGrowth$len)
hist(ToothGrowth$dose)
t.test(ToothGrowth$len)
```
#Both show a p-value of less than 2.2e-16!

#4. Figure out which transformation could make it normal
#Transform variables as necessary.
```{r}
log_airquality=log(airquality)

log_ToothGrowth=log(ToothGrowth$len)
```


#6. Make scatterplots of all the non-categorical variables
```{r}
plot(ToothGrowth$dose, ToothGrowth$len)
```

```{r}
plot(airquality$Ozone, airquality$Solar.R)
```

```{r}
plot(airquality$Ozone, airquality$Wind)
```

```{r}
plot(airquality$Temp, airquality$Month)
```


#7. Test all pairs of correlations
```{r}
cor.test(ToothGrowth$dose, ToothGrowth$len)
#p-value of 1.233e-14
```

```{r}
cor.test(airquality$Ozone, airquality$Solar.R)
#p-value of 0.0001793
```

```{r}
cor.test(airquality$Ozone, airquality$Wind)
#p-value of 9.272e-13
```

```{r}
cor.test(airquality$Temp, airquality$Month)
#p-value of 6.026e-08
```

#8. Adjust p-value using fdr and Bonferroni
```{r}
pvalues = c(1.233e-14, 0.0001793, 9.272e-13, 6.026e-08)
adjusted_pvalues <- p.adjust(pvalues, method="BH")
print(adjusted_pvalues)

```


#Specific tests
#TootGrowth:
#Perform an ANOVA with supp as the factor.
```{r}
tg_anova = oneway.test(dose~supp) #where the second portion of the parentheses is the factor.
tg_anova

```

#Perform a linear regression of len with dose
```{r}
fit = lm(len~dose, data=ToothGrowth)
plot(fit)
```

#Airquality:
#Perform an ANOVA of temp with month as factor
#Correlate Ozone with Solar.R, Wind with Temp
```{r}
anov = oneway.test(Temp~Month, data=airquality)
anov
cor.test(airquality$Ozone, airquality$Solar.R)
cor.test(airquality$Wind, airquality$Temp)
